---
title: "project7_ml"
author: "Jeff Tjeuw"
date: "11/06/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We first take the enron financial data that we extracted as a csv.

List here of the features in the csv.
extract_features = ["poi",
                    "salary",
                    "deferral_payments",
                    "total_payments",
                    "loan_advances",
                    "bonus",
                    "restricted_stock_deferred",
                    "deferred_income",
                    "total_stock_value",
                    "expenses",
                    "exercised_stock_options",
                    "other",
                    "long_term_incentive",
                    "restricted_stock",
                    "director_fees"]

```{r import_data, echo=FALSE}
setwd(
  "~/data_research/u_machine_learning/ud120-projects/final_project")
enron_fin_data <- read.csv(
  "enron_financial_features.csv", header=FALSE)
my_headings <- c("poi",
                 "salary",
                 "deferral_payments",
                 "total_payments",
                 "loan_advances",
                 "bonus",
                 "restricted_stock_deferred",
                 "deferred_income",
                 "total_stock_value",
                 "expenses",
                 "exercised_stock_options",
                 "other",
                 "long_term_incentive",
                 "restricted_stock",
                 "director_fees")
library(ggplot2)
names(enron_fin_data) <- my_headings
head(enron_fin_data)
# Remove the totals row from our data
tot_r = which.max(enron_fin_data$salary)
enron_fin_data <- enron_fin_data[-c(tot_r),]
```

## Function to Plot 2 Variables with POIs Shown
Create a function to plot POIs - to save time
```{r plot_function, echo=FALSE}
plot_pairs <- function(var1, var2){
  ggplot(enron_fin_data, aes_string(x = var1, y = var2)) +
    geom_point(aes(color = factor(poi))) +
    scale_color_manual(values = c("black", "red"))
}
```
Generate a selection of plots
```{r create_plots, echo=FALSE}
plot_pairs("salary", "bonus")
```
* Neither value appears to give much information on its own, however taken as a ratio this may provide more information.

```{r}
plot_pairs("deferral_payments", "total_payments")
```
* POIs seem to have almost no deferral_payments - this could be a hint that there is a POI.

```{r}
plot_pairs("total_payments", "loan_advances")
```
* Ken Lay's outrageous loans stands out, however nothing else for the rest of the POIs.
* For the next graph however it may be useful to restrict the axes.

```{r}
plot_pairs("restricted_stock_deferred", "deferred_income")
```
* No clear relationship stands out, however it does look like we have one bad data point - the positive value for restricted_stock_deferred. Looking into the original data - all the values are negative. Interestingly no POI had any restricted_stock_deferred.

```{r}
plot_pairs("total_stock_value", "expenses")
```
* No clear relationships stand out.

```{r}
plot_pairs("exercised_stock_options", "other")
```
* There does not appear to be any clear relationships here.

```{r}
plot_pairs("long_term_incentive", "restricted_stock")
```
* In the graph of restricted_stock vs long_term_incentive hints at a possible relationship.The red POI points seem to line up on an axis. Initially this could make sense as persons with a larger proportion of their compensation tied up in restricted_stock have an incentive to keep the fraud going.

```{r}
plot_pairs("total_payments", "director_fees")
```
* From the last graph we can straight away rule out director_fees as a field for identifying POIs. It appears most of the independent directors that received fees were not involved in the fraud perpetuated.

## Features for Further Investigation
* restricted_stock
* long_term_incentive
* salary
* bonus
* exercised_stock_options
* total_stock_value
* deferred_income
* total_payments
This is after ruling out a few that seem unlikely to be of interest. The next approach might be to take the absolute skew out of the numbers (something that we saw worked well when dealing with emails) and take values as a ratio - this will tend to but the numbers into similar orders of magnitude.

Given that the fraud involved creating ficticious and inflated asset values to keep the company "afloat", one could assume that the POIs engaged in the fraud would have been aware of this. Therefore they would have known that the company did not have sufficient cash. Therefore they would have eschewed cash payments in exchange for stock. We test this theory out by creating a ratio of total_stock_value to total_payments ratio.

```{r}
enron_fin_data$stock_payment_ratio <- 
  enron_fin_data$total_stock_value / enron_fin_data$total_payments
ggplot(enron_fin_data, aes(x = stock_payment_ratio)) +
  geom_dotplot(aes(color = factor(poi))) +
  scale_color_manual(values = c("black", "red"))
```

```{r}
plot_pairs("stock_payment_ratio", "total_payments")
```
Initial plots don't look promising. However we see that this could be blown out by extremely small denominators. We investigate what it looks like when we investigate stock payments as a percentage of total payments.
```{r}
enron_fin_data$stock_percentage <- enron_fin_data$total_stock_value / 
  (enron_fin_data$total_stock_value + enron_fin_data$total_payments)
plot_pairs("stock_percentage", "total_payments")
```
Again the results do not look very promising.

We go back to some of the other visually identified ratios:
```{r}
enron_fin_data$rs_lti_ratio <- enron_fin_data$restricted_stock /
  enron_fin_data$long_term_incentive
plot_pairs("rs_lti_ratio", " total_payments")
```
```{r}
enron_fin_data$deferral_ratio <- enron_fin_data$deferral_payments / 
  enron_fin_data$total_payments
plot_pairs("deferral_ratio", "rs_lti_ratio")
```
```{r}
plot_pairs("exercised_stock_options", "bonus")
```

```{r}
enron_fin_data$bonus_percentage <- enron_fin_data$bonus / 
  (enron_fin_data$bonus + enron_fin_data$salary)
enron_fin_data$exe_opt_percentage <- enron_fin_data$exercised_stock_options / 
  (enron_fin_data$total_stock_value)
plot_pairs("bonus_percentage", "exe_opt_percentage")
```

```{r}
plot_pairs("bonus_percentage", "stock_percentage")
```
Here we start to see a possible cluster emerge - the red dots are largely concentrated to the right of the graph.

Maybe we consider non-financial payments as well.
```{r}
enron_fin_data$non_salary_percent <-
  1 - enron_fin_data$salary / enron_fin_data$total_payments
plot_pairs("non_salary_percent", "bonus_percentage")
```
```{r}
plot_pairs("non_salary_percent", "exe_opt_percentage")
```
```{r}
plot_pairs("non_salary_percent", "total_stock_value")
```
```{r}
enron_fin_data$stock_salary_ratio <-
  enron_fin_data$total_stock_value / enron_fin_data$salary
plot_pairs("stock_salary_ratio", "non_salary_percent") +
  scale_x_continuous(limits = c(0, 150))
```
POIs have high non-salary percentages, but so do many people.
```{r}
plot_pairs("non_salary_percent", "total_stock_value")
```
```{r}
enron_fin_data$non_salary_payments <-
  enron_fin_data$total_payments - enron_fin_data$salary
plot_pairs("non_salary_payments", "total_stock_value") +
  scale_x_continuous(limits = c(0, 12500000)) +
  scale_y_continuous(limits = c(0, 12500000))
```
Focus on a second dimension in addtion to non_salary_percent

```{r}
enron_fin_data$bs_ratio <- enron_fin_data$bonus/enron_fin_data$salary
plot_pairs("exercised_stock_options", "bs_ratio")
```

